---
title: "Labs Overview: Getting to know _R_ and _Bioconductor_"
author: "Valerie Obenchain (valerie.obenchain@roswellpark.org)<br />
    Lori Shepherd (lori.shepherd@roswellpark.org)<br />
    Martin Morgan (martin.morgan@roswellpark.org)<br />
    Stanford University, Stanford, CA<br />
    25 - 26 June, 2016"
output:
  BiocStyle::html_document:
    toc: true
vignette: >
  % \VignetteIndexEntry{Labs Overview}
  % \VignetteEngine{knitr::rmarkdown}
---

```{r style, echo = FALSE, results = 'asis'}
BiocStyle::markdown()
```

```{r setup, echo=FALSE}
knitr::opts_chunk$set(cache=TRUE)
```

This lab has four parts. Each part consists of informative material
followed by short or intermediate length exercises and case
studies. The material introduces or refreshes your knowledge of _R_,
and walks through some of the central classes and annotation resources
for working with high-throughput sequence data in _Bioconductor_.

It's recommended that you choose two of the following sections, and
spend about 20 minutes on each. Take time to work through the code in
the material as well as the exercises, asking questions of your
colleagues and the TAs as needed.

1 [Introduction to _R_](L1-r-intro.html)

2 [Introduction to _Bioconductor_](L2-bioc-intro.html)

3 [Data Representations](L3-bioc-data-representation.html)

4 [Annotation](L4-bioc-annotation.html)
